Constraint Grammars as a SAT problem

### 2015-03-31 paper/nodalida2015.tex

\title{Constraint Grammar as a SAT problem}

@@

We represent Constraint Grammar \cite{karlsson1995constraint} 
as a Boolean satisfiability (SAT) problem.
This is attractive from several reasons: formal logic is
well-studied, and serves as an abstract language to reason about the
properties of CG. Constraint rules encoded in logic capture richer
dependencies between the tags than standard CG.

@@

SAT solver performs a search, 
and starts building possible models that satisfy both constraints. 
In addition to the given constraints, we have default rules to emulate
the CG principles: an analysis is true if no rule affects it,
and at least one analysis for each word is true---the notion of
``last'' is not applicable.

@@

\noindent These three features might change the way that rules are written. 
We predict that less rules are needed; whether this holds in
the order of thousands of rules remains to be tested.
Getting rid of ordering could ease the task of the grammar writer,
since it removes the need to estimate the best sequence of rule
applications.

However, the implementation should not be too different from the
existing ones---our preliminary evaluation against VISL CG-3 shows
promising results in this regard.
\pagebreak

@@

The worst-case complexity of SAT is exponential, whereas the standard
implementations of CG are polynomial, but with advances in SAT solving
techniques, the performance in the average case is much more feasible than in the previous works done in 90s--00s.
We used the open-source SAT solver MiniSat \cite{een04sat}.

@@

>>>>

\title{Constraint Grammars as a SAT problem}

@@

We represent Constraint Grammars (CG) \cite{karlsson1995constraint} 
as a Boolean satisfiability (SAT) problem.
This is attractive from several reasons: formal logic is
well-studied, and serves as an abstract language to reason about the
properties of CG. Constraint rules encoded in logic capture richer
dependencies between the tags than standard CG.

@@

The SAT solver performs a search, 
and starts building possible models that satisfy both constraints. 
In addition to the given constraints, we have default rules to emulate
the CG principles: an analysis is true if no rule affects it,
and at least one analysis for each word is true---the notion of
``last'' is not applicable.

@@

\noindent These three features influence the way that rules are written. 
We predict that less rules are needed; whether this holds in
the order of thousands of rules remains to be tested.
Getting rid of ordering could ease the task of the grammar writer,
since it removes the need to estimate the best sequence of rule
applications.

However, the results using our approach should not be too different from the
existing ones---our preliminary evaluation against VISL CG-3 shows
promising results in this regard.
\pagebreak

@@

The worst-case complexity of SAT is exponential, whereas the standard
implementations of CG are polynomial, but with advances in SAT solving
techniques, the performance in the average case in practice is much more feasible than in the previous works done in 90s--00s.
We used the open-source SAT solver MiniSat \cite{een04sat}.

@@

\section{Conclusions}

SAT-solvers are nowadays powerful enough to be used for dealing with Constraint Grammars. A logic-based approach to CG has possible advantages over more traditional approaches; a SAT-solver may disambiguate more words, and may do so more precisely. Also, the SAT-solver requires less rules, and these rules are simpler. The ordering of rules is something we found was incompatible with a logic-based approach. We compensate this by maximising rule applications. Whether this is an advantage or disadvantage remains to be seen. Our initial experimental results are promising.
